Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1]}

### 2017-12-14 paper/paper.tex
More writing

@@

The subgraph isomorphism problem is to find a copy of a small ``pattern'' graph inside a larger
``target'' graph.  The current single strongest subgraph isomorphism solver uses ``highest degree
first'' as a value-ordering heuristic to direct a constraint programming style search
\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
heuristic is far from perfect. Furthermore, even if degree were a perfect source of information,
many target graphs do not have a large degree distribution, so we would still need to determine
\emph{which} vertex of highest degree would be best.

@@

\section{The Basic Algorithm}

?? Non-induced!

?? Notation: $\vertexset(G)$, $\neighbourhood_G(v)$, $\nds_G(v)$, $\preceq$, $\sim_G$, $G^{a,b}$, $\mapsto$.

?? Started with the Glasgow algorithm, removed parts which did not contribute hugely to speedups,
and ended up with a simpler algorithm that performs as well in practice. In particular, because we
will be working with reasonably large target graphs, no ILF, and not doing supplemental graphs of
length 3. This is \cref{algorithm:sip}, ignoring the shaded parts, which are introduced in the
context of restarts.

?? Briefly describe this, probably referring elsewhere as far as possible to save space.

\newcommand{\siplabel}[1]{\label{line:sip:#1}}
\newcommand{\siplineref}[1]{line~\ref{line:sip:#1}}
\begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
    \begin{tikzpicture}[remember picture,overlay]
        \coordinate (restart1sc) at ($(pic cs:restart1s) + (0, 0.15)$);
        \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart1sc) (restart1ec)] { };
        \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
        \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart2sc) (restart2ec)] { };
        \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
        \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
        \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
        \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
        \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
        \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
        \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
        \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
        \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec)] { };
        \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
        \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
        \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
        \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
        \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
        \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart9sc) (restart9ec)] { };
    \end{tikzpicture}

@@

    \nl \ForEach{$v \in \vertexset(\mathcal{P})$}{

@@

            \\\hspace*{4em}$\nds_P(v) \preceq \nds_T(w)\}$ \;

@@

        \nl \If{\FuncSty{propagate}(D')}{
            \nl $S \gets \FuncSty{search}(\{ E \in D' : \left|E\right| > 1 \}, B \wedge (v \mapsto v'))$ \;

@@

    \nl $\KwSty{return}~\KwSty{false}$

@@

\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$ \;

@@

    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}

@@

\caption{Non-induced subgraph isomorphism. Shaded code is added for the restarting variant.}

>>>>

\newcommand{\siplabel}[1]{\label{line:sip:#1}}
\newcommand{\siplineref}[1]{line~\ref{line:sip:#1}}
\newcommand{\siplinerangeref}[2]{lines~\ref{line:sip:#1} to~\ref{line:sip:#2}}

@@

The subgraph isomorphism problem is to find a copy of a small ``pattern'' graph inside a larger
``target'' graph.  The current single strongest subgraph isomorphism solver uses ``highest degree
first'' as a value-ordering heuristic to direct a constraint programming style search
\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:conf/ijcai/McCreeshPT16}. This
heuristic is far from perfect---and even if degree were a perfect source of information,
many target graphs do not have a large degree distribution, so we would still need to determine
\emph{which} vertex of highest degree would be best.

@@

\section{Definitions and Algorithms}

Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
that vertices $v$ and $w$ are adjacent. The \emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$,
is the set of vertices adjacent to $v$, and the \emph{degree} of $v$, $\deg_G(v)$, is the
cardinality of its neighbourhood; when the graph is clear from the context, we omit the $G$
subscript. The \emph{neighbourhood degree sequence} of $v$, $\nds_G(v)$, is the sequence of degrees
of the neighbours of $v$ in descending order, and we denote by $\preceq$ the inclusion relation of
\citet{DBLP:journals/constraints/ZampelliDS10}. As per \citet{DBLP:conf/aaai/HoffmannMR17},
$G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$, and with edges between vertices $v$
and $v$ if there are at least $n$ simple paths of length exactly $\ell$ between vertices $v$ and $w$
in $G$.

The non-induced subgraph isomorphism problem is to find an injective mapping from the vertices of a
pattern graph $\mathcal{P}$ to a target graph $\mathcal{T}$, such that adjacent vertices in
$\mathcal{P}$ are mapped to adjacent vertices in $\mathcal{T}$. We write $v \mapsto v'$ to mean that
pattern vertex $v$ is mapped to target $v'$ under such a mapping.

\subsection{A Subgraph Isomorphism Algorithm}

We now briefly outline the non-shaded parts of \cref{algorithm:sip}, which is how we will solve the
subgraph isomorphism problem. This algorithm is very closely based upon the $k{\downarrow}$
algorithm of \citet{DBLP:conf/aaai/HoffmannMR17} with $k = 0$, and we refer the reader to that paper
for full technical details; that algorithm, in turn, is a simplification of the Glasgow2 algorithm
\citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}.

\Cref{algorithm:sip} works by recursively building up a valid mapping of pattern vertices to target
vertices. For each pattern vertex $v$, we have a domain $D_v$ containing a set of feasible target
vertices. Each domain is initialised at the top of search (\siplinerangeref{initstart}{initend}),
using the invariants described by \citet{DBLP:conf/aaai/HoffmannMR17} to restrict its initial
values.

The $\FuncSty{search}$ procedure recursively selects an unassigned domain $D$
(\siplineref{variableordering}), and tries assigning
it each of its remaining value in turn (\siplineref{valueordering}). The effects of this assignment
are propagated (\siplineref{dopropagate}). If an
inconsistency is detected, we continue and try another value; otherwise, we recurse
(\siplineref{recurse}) until a solution is found. Finally, if we run out of values, we backtrack
(\siplineref{backtrack}).

Our $\FuncSty{propagate}$ routine is similarly close to that of \citet{DBLP:conf/aaai/HoffmannMR17},
differing only in that we carry out the filtering for injectivity
(\siplinerangeref{alldiffstart}{alldiffend}) after every assignment rather than once at the end of
the unit propagation loop. (We make this change because the all-different propagator used is not
idempotent, and we wish to avoid an odd interaction when we introduce nogood filtering.)

\begin{algorithm}[p]\DontPrintSemicolon\small\SetInd{0.1em}{1em}
    \begin{tikzpicture}[remember picture,overlay]
        \coordinate (restart1sc) at ($(pic cs:restart1s) + (0, 0.15)$);
        \coordinate (restart1ec) at ($(pic cs:restart1e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart1sc) (restart1ec)] { };
        \coordinate (restart2sc) at ($(pic cs:restart2s) + (0, 0.15)$);
        \coordinate (restart2ec) at ($(pic cs:restart2e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart2sc) (restart2ec)] { };
        \coordinate (restart3sc) at ($(pic cs:restart3s) + (0, 0.15)$);
        \coordinate (restart3ec) at ($(pic cs:restart3e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart3sc) (restart3ec)] { };
        \coordinate (restart4sc) at ($(pic cs:restart4s) + (0, 0.15)$);
        \coordinate (restart4ec) at ($(pic cs:restart4e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart4sc) (restart4ec)] { };
        \coordinate (restart5sc) at ($(pic cs:restart5s) + (0, 0.15)$);
        \coordinate (restart5ec) at ($(pic cs:restart5e) - (0, 0.10)$);
        \coordinate (restart5rc) at ($(pic cs:restart5r) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart5sc) (restart5ec) (restart5rc)] { };
        \coordinate (restart6sc) at ($(pic cs:restart6s) + (0, 0.15)$);
        \coordinate (restart6ec) at ($(pic cs:restart6e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart6sc) (restart6ec)] { };
        \coordinate (restart7sc) at ($(pic cs:restart7s) + (0, 0.15)$);
        \coordinate (restart7ec) at ($(pic cs:restart7e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart7sc) (restart7ec)] { };
        \coordinate (restart8sc) at ($(pic cs:restart8s) + (0.05, 0.15)$);
        \coordinate (restart8ec) at ($(pic cs:restart8e) + (0, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart8sc) (restart8ec)] { };
        \coordinate (restart9sc) at ($(pic cs:restart9s) + (0, 0.15)$);
        \coordinate (restart9ec) at ($(pic cs:restart9e) + (-0.15, 0.00)$);
        \node [fill=highlight, rounded corners, fit=(restart9sc) (restart9ec)] { };
    \end{tikzpicture}

@@

    \nl \ForEach{\siplabel{initstart}$v \in \vertexset(\mathcal{P})$}{

@@

            \\\hspace*{4em}$\nds_P(v) \preceq \nds_T(w)\}$\siplabel{initend} \;

@@

        \nl \If{\FuncSty{propagate}(D')\siplabel{dopropagate}}{
            \nl $S \gets \FuncSty{search}(\{ E \in D'\,{:}\,\left|E\right|{>}\,1 \}, B \wedge (v\,{\mapsto}\,v'))$\siplabel{recurse} \;

@@

    \nl $\KwSty{return}~\KwSty{false}$\siplabel{backtrack}

@@

\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,0)$\siplabel{alldiffstart} \;

@@

    \nl \lIf{$|A| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\siplabel{hall}}\siplabel{alldiffend}

@@

\caption{Non-induced subgraph isomorphism. The shaded code is added for the restarting variant.}
